---
title: "PURM Document"
output: html_document
---

## Setting Directory

We first set the directory for the EPA plots. Installing required packages.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/karansampath/Desktop/Research/PURM')

# Put in message = false, warning = false
```


## Installing Packages

```{r eval=FALSE, include=FALSE}
if (!require(dplyr)) {
    install.packages("dplyr", repos = "http://cran.us.r-project.org")
    require(dplyr)
}
if (!require(sf)) {
    install.packages("sf", repos = "http://cran.us.r-project.org")
    require(sf)
}
if (!require(rgdal)) {
    install.packages("rgdal", repos = "http://cran.us.r-project.org")
    require(rgdal)
}

if (!require(raster)) {
    install.packages("raster", repos = "http://cran.us.r-project.org")
    require(raster)
}
if (!require(tigris)) {
    install.packages("tigris", repos = "http://cran.us.r-project.org")
    require(tigris)
}
if (!require(tmap)) {
    install.packages("tmap", repos = "http://cran.us.r-project.org")
    require(tmap)
}
if (!require(SpatialPack)) {
    install.packages("SpatialPack", repos = "http://cran.us.r-project.org")
    require(SpatialPack)
}
if(!require(ggcorrplot)){
    install.packages("ggcorrplot", repos = "http://cran.us.r-project.org")
    require(ggcorrplot)
}
if(!require(psych)){
    install.packages("psych", repos = "http://cran.us.r-project.org")
    require(psych)
}
if(!require(ggpubr)){
    install.packages("ggpubr", repos = "http://cran.us.r-project.org")
    require(ggpubr)
}
if(!require(Hmisc)){
    install.packages("Hmisc", repos = "http://cran.us.r-project.org")
    require(Hmisc)
}
if(!require(corrplot)){
    install.packages("corrplot", repos = "http://cran.us.r-project.org")
    require(corrplot)
}
```

## Loading Libraries

We load all required libraries together.

```{r libraries,  message=FALSE}
library(sf)
library(corrr)
library(tmap)
library(rgdal)
library(dplyr)
library(raster)
library(data.table)
library(sp)
library(tidyverse)
library(tidycensus)
library(psych)
library(tigris)
library(leaflet)
library(maps)
library(gstat)
library(chron)
library(maptools)
library(ggcorrplot)
library(ggpubr)
library(Hmisc)
library(corrplot)
```

##Inputting CSVs

We inputting the various csv files of different pollutants. (What are they, where they came from, what all they did, what's the goal, etc.). (Show fancier tables, like cable)

```{r input}
#read in and wrangle epa data for SO2 
so2_epa <- read.csv("SO2.csv")
so2_epa_df <- so2_epa %>% 
  dplyr::mutate(
    Date = as.factor(gsub("/00","/",strftime(as.Date(Date, format="%m/%d/%Y"), format="%m/%d/%Y"))),
    Day = gsub("/([^/]*)$","",Date),
    Year = as.factor(paste0("20",gsub(".*/","",Date))),
    Mean_value = round(SO2,2),
    SiteName = as.factor(gsub(".*[(]","(","Site Name")),
    Parameter_name = as.character("SO2")) %>% dplyr::rename(Latitude = SITE_LATITUDE, Longitude = SITE_LONGITUDE)


so2_epa_df <- so2_epa_df %>% dplyr::filter(COUNTY == "Philadelphia")

#read in and wrangle epa data for PM2.5 
pm_epa <- read.csv("PM25.csv")

pm_epa_df <- pm_epa %>% 
  dplyr::mutate(
    Date = as.factor(gsub("/00","/",strftime(as.Date(Date, format="%m/%d/%Y"), format="%m/%d/%Y"))),
    Day = gsub("/([^/]*)$","",Date),
    Year = as.factor(paste0("20",gsub(".*/","",Date))),
    Mean_value = round(Daily.Mean.PM2.5.Concentration,2),
    SiteName = as.factor(gsub(".*[(]","(","Site Name")),
    Parameter_name = as.character("PM2.5")) %>% dplyr::rename(Latitude = SITE_LATITUDE, Longitude = SITE_LONGITUDE)


pm_epa_df <- pm_epa_df %>% dplyr::filter(COUNTY == "Philadelphia")

#read in and wrangle epa data for Ozone 
pm_oz <- read.csv("Ozone.csv")

pm_oz_df <- pm_oz %>% 
  dplyr::mutate(
    Date = as.factor(gsub("/00","/",strftime(as.Date(Date, format="%m/%d/%Y"), format="%m/%d/%Y"))),
    Day = gsub("/([^/]*)$","",Date),
    Year = as.factor(paste0("20",gsub(".*/","",Date))),
    Mean_value = round(Daily.Max.8.hour.Ozone.Concentration,2),
    SiteName = as.factor(gsub(".*[(]","(","Site Name")),
    Parameter_name = as.character("OZ")) %>% dplyr::rename(Latitude = SITE_LATITUDE, Longitude = SITE_LONGITUDE)


pm_oz_df <- pm_oz_df %>% dplyr::filter(COUNTY == "Philadelphia")


#read in and wrangle epa data for NO2 
pm_no2 <- read.csv("NO2.csv")

pm_no2_df <- pm_no2 %>% 
  dplyr::mutate(
    Date = as.factor(gsub("/00","/",strftime(as.Date(Date, format="%m/%d/%Y"), format="%m/%d/%Y"))),
    Day = gsub("/([^/]*)$","",Date),
    Year = as.factor(paste0("20",gsub(".*/","",Date))),
    Mean_value = round(Daily.Max.1.hour.NO2.Concentration,2),
    SiteName = as.factor(gsub(".*[(]","(","Site Name")),
    Parameter_name = as.character("NO2")) %>% dplyr::rename(Latitude = SITE_LATITUDE, Longitude = SITE_LONGITUDE)


pm_no2_df <- pm_no2_df %>% dplyr::filter(COUNTY == "Philadelphia")


#read in and wrangle epa data for Lead 
pm_pb <- read.csv("Pb.csv")

pm_pb_df <- pm_pb %>% 
  dplyr::mutate(
    Date = as.factor(gsub("/00","/",strftime(as.Date(Date, format="%m/%d/%Y"), format="%m/%d/%Y"))),
    Day = gsub("/([^/]*)$","",Date),
    Year = as.factor(paste0("20",gsub(".*/","",Date))),
    Mean_value = round(Daily.Mean.Pb.Concentration,2),
    SiteName = as.factor(gsub(".*[(]","(","Site Name")),
    Parameter_name = as.character("Pb")) %>% dplyr::rename(Latitude = SITE_LATITUDE, Longitude = SITE_LONGITUDE)


pm_pb_df <- pm_pb_df %>% dplyr::filter(COUNTY == "Philadelphia")

#read in and wrangle epa data for CO 
pm_co <- read.csv("CO.csv")

pm_co_df <- pm_co %>% 
  dplyr::mutate(
    Date = as.factor(gsub("/00","/",strftime(as.Date(Date, format="%m/%d/%Y"), format="%m/%d/%Y"))),
    Day = gsub("/([^/]*)$","",Date),
    Year = as.factor(paste0("20",gsub(".*/","",Date))),
    Mean_value = round(Daily.Max.8.hour.CO.Concentration,2),
    SiteName = as.factor(gsub(".*[(]","(","Site Name")),
    Parameter_name = as.character("CO")) %>% dplyr::rename(Latitude = SITE_LATITUDE, Longitude = SITE_LONGITUDE)


pm_co_df <- pm_co_df %>% dplyr::filter(COUNTY == "Philadelphia")


#read in and wrangle epa data for PM10 
pm_10 <- read.csv("PM10.csv")

pm_10_df <- pm_10 %>% 
  dplyr::mutate(
    Date = as.factor(gsub("/00","/",strftime(as.Date(Date, format="%m/%d/%Y"), format="%m/%d/%Y"))),
    Day = gsub("/([^/]*)$","",Date),
    Year = as.factor(paste0("20",gsub(".*/","",Date))),
    Mean_value = round(Daily.Mean.PM10.Concentration,2),
    SiteName = as.factor(gsub(".*[(]","(","Site Name")),
    Parameter_name = as.character("PM10")) %>% dplyr::rename(Latitude = SITE_LATITUDE, Longitude = SITE_LONGITUDE)


pm_10_df <- pm_10_df %>% dplyr::filter(COUNTY == "Philadelphia")

head(pm_10_df)
```

## Creating Base Raster

```{r work}
## 0.1km Rasters for Philadelphia only
getRaster_ph <- function(data, par_name){
  ## filter data to contiguous US
  ## get one row per location by averaging all entries
  dat <- dplyr::filter(data,
                       Parameter_name == par_name) %>%
    dplyr::filter(!is.na(Mean_value), Mean_value >= 0) %>%
    dplyr::group_by(Latitude, Longitude) %>%
    dplyr::summarise(avg = mean(Mean_value, na.rm = TRUE))
  
  if(nrow(dat) > 0){
    coordinates(dat) = ~Longitude+Latitude
    crs(dat) <- CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
    
    ## make base raster
    #find distance between the latitude and longitudes and convert to km (*111 for 1km)
    r <- raster(nrow = 301, ncol = 360, extent(-75.28027, -74.95576, 39.867, 40.13799)) #0.1km
    crs(r) <- CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
    ## generate raster (idw with 5 nearest sites)
    gs <- gstat(formula=avg~1, data=dat, nmax = 5)
    nn <- interpolate(r, gs)
    
    return(nn)
  } else{
    print("zero")
    #find distance between the latitude and longitudes and convert to km (*111 for 1km)
    r <- raster(nrow = 301, ncol = 360, extent(-75.28027, -74.95576, 39.867, 40.13799))
    crs(r) <- CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
    values(r) <- rep(-1, 108360)
    return(r)
  }
}

```

## Writing Rasters
```{r pm 2.5}

#PH shape file
ph_shapefile <- readOGR("./Neighborhoods_Philadelphia/Neighborhoods_Philadelphia.shp")
ph <- spTransform(x = ph_shapefile, CRSobj = '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0')

#Census Shapefile
census_shapefile <- readOGR("./census_shapefile/tl_2020_42_tract.shp")


#make pm.25 raster
pm_ph_daily_brick <- brick(getRaster_ph(pm_epa_df,"PM2.5"))
pm_ph_daily <- getRaster_ph(pm_epa_df,"PM2.5")
pm_ph_daily_brick[pm_ph_daily_brick == -1] <- NA
names(pm_ph_daily_brick) <- c("pm2.5")
c <- crop(pm_ph_daily_brick, extent(ph))
pm25_raster <- raster::mask(c, ph) 
writeRaster(pm25_raster,"ph_pm_daily_0.1km_brick.tif",overwrite=TRUE)

#make so2 raster
SO2_ph_daily_brick <- brick(getRaster_ph(so2_epa_df,"SO2"))
SO2_ph_daily_brick[SO2_ph_daily_brick == -1] <- NA
names(SO2_ph_daily_brick) <- c("so2")
#crop by shape
c <- crop(SO2_ph_daily_brick, extent(ph))
so2_raster <- raster::mask(c, ph) 
writeRaster(so2_raster,"ph_SO2_daily_0.1km_brick.tif",overwrite=TRUE)


#make CO raster
CO_ph_daily_brick <- brick(getRaster_ph(pm_co_df,"CO"))
CO_ph_daily_brick[CO_ph_daily_brick == -1] <- NA
names(CO_ph_daily_brick) <- c("CO")
#crop by shape
c <- crop(CO_ph_daily_brick, extent(ph))
co_raster <- raster::mask(c, ph) 
writeRaster(co_raster,"ph_CO_daily_0.1km_brick.tif",overwrite=TRUE)

#make Ozone raster
oz_ph_daily_brick <- brick(getRaster_ph(pm_oz_df,"OZ"))
oz_ph_daily_brick[oz_ph_daily_brick == -1] <- NA
names(oz_ph_daily_brick) <- c("OZ")
#crop by shape
c <- crop(oz_ph_daily_brick, extent(ph))
o3_raster <- raster::mask(c, ph) 
writeRaster(o3_raster,"ph_oz_daily_0.1km_brick.tif",overwrite=TRUE)

#make NO2 raster
NO2_ph_daily_brick <- brick(getRaster_ph(pm_no2_df,"NO2"))
NO2_ph_daily_brick[NO2_ph_daily_brick == -1] <- NA
names(NO2_ph_daily_brick) <- c("NO2")
#crop by shape
c <- crop(NO2_ph_daily_brick, extent(ph))
no2_raster <- raster::mask(c, ph) 
writeRaster(no2_raster,"ph_NO2_daily_0.1km_brick.tif",overwrite=TRUE)



#make Pb raster
pb_ph_daily_brick <- brick(getRaster_ph(pm_pb_df,"Pb"))
pb_ph_daily_brick[pb_ph_daily_brick == -1] <- NA
names(pb_ph_daily_brick) <- c("Pb")
#crop by shape
c <- crop(pb_ph_daily_brick, extent(ph))
pb_raster <- raster::mask(c, ph) 
writeRaster(pb_raster,"ph_pb_daily_0.1km_brick.tif",overwrite=TRUE)


#make PM10 raster
pm10_ph_daily_brick <- brick(getRaster_ph(pm_10_df,"PM10"))
pm10_ph_daily_brick[pm10_ph_daily_brick == -1] <- NA
names(pm10_ph_daily_brick) <- c("PM10")
#crop by shape
c <- crop(pm10_ph_daily_brick, extent(ph))
pm10_raster <- raster::mask(c, ph) 
writeRaster(pm10_raster,"ph_pm10_daily_0.1km_brick.tif",overwrite=TRUE)
```

```{r census means}

pm25_poly <- raster::extract(pm25_raster, census_shapefile)
pm10_poly <- raster::extract(pm10_raster, census_shapefile)
no2_poly <- raster::extract(no2_raster, census_shapefile)
o3_poly <- raster::extract(o3_raster, census_shapefile)
co_poly <- raster::extract(co_raster, census_shapefile)
so2_poly <- raster::extract(so2_raster, census_shapefile)
pb_poly <- raster::extract(pb_raster, census_shapefile)

# Note output is a list of 286 vectors, corresponding to the 286 polygons in the sepa zip/muni shapefile. Each vector contains all the raster values falling within a polygon.
str(census_shapefile, max.level = 2)
str(pm25_poly)
str(pm10_poly)
str(no2_poly)
str(o3_poly)
str(co_poly)
str(so2_poly)
str(pb_poly)

# Find the mean for each polygon
pm25_means <- unlist(lapply(pm25_poly, function(x) if (!is.null(x)) mean(x, na.rm=TRUE) else NA ))
pm10_means <- unlist(lapply(pm10_poly, function(x) if (!is.null(x)) mean(x, na.rm=TRUE) else NA ))
no2_means <- unlist(lapply(no2_poly, function(x) if (!is.null(x)) mean(x, na.rm=TRUE) else NA ))
o3_means <- unlist(lapply(o3_poly, function(x) if (!is.null(x)) mean(x, na.rm=TRUE) else NA ))
co_means <- unlist(lapply(co_poly, function(x) if (!is.null(x)) mean(x, na.rm=TRUE) else NA ))
so2_means <- unlist(lapply(so2_poly, function(x) if (!is.null(x)) mean(x, na.rm=TRUE) else NA ))
pb_means <- unlist(lapply(pb_poly, function(x) if (!is.null(x)) mean(x, na.rm=TRUE) else NA ))


# Append mean pm to SEPA data
census_shapefile@data$pm25 <- pm25_means
census_shapefile@data$pm10 <- pm10_means
census_shapefile@data$no2 <- no2_means
census_shapefile@data$o3 <- o3_means
census_shapefile@data$co <- co_means
census_shapefile@data$so2 <- so2_means
census_shapefile@data$pb <- pb_means
df_census <- census_shapefile@data
census_final <- df_census %>% dplyr::select(GEOID, pm25, pm10, no2, o3, co, so2, pb) %>% drop_na()
head(census_final)
```

```{r labelformat}
myLabelFormat = function(prefix = "", suffix = "", between = " &ndash; ", digits = 3, 
                         big.mark = ",", transform = identity, t.val = Inf) {
  formatNum <- function(x) {
    format(round(transform(x), digits), trim = TRUE, scientific = FALSE, 
           big.mark = big.mark)
  }
  function(type, ...) {
    switch(type, numeric = (function(cuts) {
      cuts <- sort(cuts, decreasing = T) #just added
      paste0(prefix, formatNum(cuts), ifelse(cuts == t.val, "+", ""))
    })(...), bin = (function(cuts) {
      n <- length(cuts)
      paste0(prefix, formatNum(cuts[-n]), between, formatNum(cuts[-1]), 
             suffix)
    })(...), quantile = (function(cuts, p) {
      n <- length(cuts)
      p <- paste0(round(p * 100), "%")
      cuts <- paste0(formatNum(cuts[-n]), between, formatNum(cuts[-1]))
      paste0("<span title=\"", cuts, "\">", prefix, p[-n], 
             between, p[-1], suffix, "</span>")
    })(...), factor = (function(cuts) {
      paste0(prefix, as.character(transform(cuts)), suffix)
    })(...))
  }
}


```


## Plotting Maps

```{r map}
make_map <- function(ras.t,title="PM2.5",limt,all_values,monitors_ph,palette){
  if (palette %in% c("heat","HEAT","h","H")){
    palette1 <- colorNumeric(rev(heat.colors(8)), all_values, na.color = "transparent") #terrain
    palette2 <- colorNumeric(heat.colors(8), all_values, na.color = "transparent")
  } else if (palette %in% c("terrain","TERRAIN","t","T")) {
    palette1 <- colorNumeric(rev(terrain.colors(8)), all_values, na.color = "transparent")
    palette2 <- colorNumeric(terrain.colors(8), all_values, na.color = "transparent")
  }
map1<-leaflet(ph) %>% 
    setView(-75.14, 40, zoom = 11) %>%
    addRasterImage(x = ras.t, colors = palette1, method = "ngb", opacity = 0.7) %>%
    addPolygons(color = "black", weight = 1, fillColor = "transparent") %>%
    addCircleMarkers(lng = monitors_ph$Longitude, lat = monitors_ph$Latitude, radius = 5,label = monitors_ph$SiteName) %>%
    addLegend(pal = palette2,
              values = all_values,
              labFormat = myLabelFormat(t.val = limt), title = title,
              position = "bottomleft")
map1
}
```

## Making the Bricks

```{r PM2.5 brick}
pm2.5.brick <- brick("ph_pm_daily_0.1km_brick.tif")
#Use max value to reclassify
pm.t1 <- reclassify(pm2.5.brick[[1]], c(11, Inf, 11)) #8
pm25_data <- values(pm.t1)
epa_sites <- pm_epa_df %>% dplyr::select(SiteName, Latitude, Longitude) %>% unique()
epa_main <- pm_epa_df %>% dplyr::group_by(Latitude, Longitude) %>% dplyr::summarise(m = mean(Mean_value))
P1 <- make_map(pm.t1,title="PM2.5",11,pm25_data,epa_sites,palette="h") 
P1
```

```{r CO brick}
pm2.5.brick <- brick("ph_CO_daily_0.1km_brick.tif")
#Use max value to reclassify
pm.t2 <- reclassify(pm2.5.brick[[1]], c(11, Inf, 11)) #8
pm25_data <- values(pm.t1)
epa_sites <- pm_co_df %>% dplyr::select(SiteName, Latitude, Longitude) %>% unique()
P2 <- make_map(pm.t1,title="CO",11,pm25_data,epa_sites,palette="h") 
P2
```

```{r NO2 brick}
pmno2.brick <- brick("ph_NO2_daily_0.1km_brick.tif")
#Use max value to reclassify
pm.t1 <- reclassify(pmno2.brick[[1]], c(11, Inf, 11)) #8
pmno2_data <- values(pm.t1)
epa_sites <- pm_no2_df %>% dplyr::select(SiteName, Latitude, Longitude) %>% unique()
P3 <- make_map(pm.t1,title="NO2",11,pmno2_data,epa_sites,palette="h")
P3
```


```{r SO2 brick}
pm2.5.brick <- brick("ph_SO2_daily_0.1km_brick.tif")
#Use max value to reclassify
pm.t1 <- reclassify(pm2.5.brick[[1]], c(11, Inf, 11)) #8
pm25_data <- values(pm.t1)
epa_sites <- so2_epa_df %>% dplyr::select(SiteName, Latitude, Longitude) %>% unique()
P3 <- make_map(pm.t1,title="SO2",11,pm25_data,epa_sites,palette="h") #save image as precovid 2020
P3
```





```{r Ozone brick}
pm2.5.brick <- brick("ph_oz_daily_0.1km_brick.tif")
#Use max value to reclassify
pm.t1 <- reclassify(pm2.5.brick[[1]], c(11, Inf, 11)) #8
pm25_data <- values(pm.t1)
epa_sites <- pm_oz_df %>% dplyr::select(SiteName, Latitude, Longitude) %>% unique()
P4 <- make_map(pm.t1,title="Ozone",11,pm25_data,epa_sites,palette="h") #save image as precovid 2020
P4
```
```{r PM10 brick}
pm2.5.brick <- brick("ph_pm10_daily_0.1km_brick.tif")
#Use max value to reclassify
pm.t1 <- reclassify(pm2.5.brick[[1]], c(11, Inf, 11)) #8
pm25_data <- values(pm.t1)
epa_sites <- pm_10_df %>% dplyr::select(SiteName, Latitude, Longitude) %>% unique()
P5 <- make_map(pm.t1,title="PM10",11,pm25_data,epa_sites,palette="h") #save image as precovid 2020
P5
```

```{r Pb brick}
pb.brick <- brick("ph_pb_daily_0.1km_brick.tif")
#Use max value to reclassify
pm.t1 <- reclassify(pb.brick[[1]], c(11, Inf, 11)) #8
pb_data <- values(pm.t1)
epa_sites <- pm_pb_df %>% dplyr::select(SiteName, Latitude, Longitude) %>% unique()
P6 <- make_map(pm.t1,title="Pb",11,pb_data,epa_sites,palette="h")
P6
## Make a table with tracts, automatically should come with GEOIDs once you convert bricks to table (need to find GEOIDs and values for the entire raster)
```

```{r together}
library(htmltools)
leaflet_grid <- 
  tagList(
    tags$table(width = "100%",
      tags$tr(
          tags$td(P1),
          tags$td(P2)
      ),
      tags$tr(
        tags$td(P3),
        tags$td(P4)
      )
    )
  )
browsable(leaflet_grid)
```
```{r raster corr}
library(gstat)                                         
library(sp)
z <- stack(pm.t1, pm.t2)
r <- calc(z, fun=function(x) cor(x[1], x[2], method='pearson'))
plot(pm.t1)
```

## Plotting Census Data

```{r census}
#census_api_key("3304af6f650f9f6e42986fbc8497d3b58ac1250e", install=TRUE, overwrite = TRUE)
v17 <- load_variables(dataset = "acs5", year = 2019)
view(v17)
```

```{r race}
race <- get_acs(geography = "tract", year = 2019, variables = c("B03002_001", "B03002_001","B03002_002","B03002_003","B03002_004","B03002_005", "B03002_006", "B03002_007","B03002_008", "B03002_009", "B03002_010", "B03002_011"), state = "PA", county = c("Philadelphia", "Chester", "Montgomery", "Bucks", "Delaware"))
rtracts <- tracts(state = 'PA', county = c("Philadelphia", "Chester" , "Montgomery", "Bucks", "Delaware"), cb = TRUE)
class(race)
class(rtracts)
plot(st_geometry(rtracts))
```

```{r hisp plot}
race_merged<- geo_join(rtracts, race, "GEOID", "GEOID")
race_merged
popup <- paste0("GEOID: ", race_merged$GEOID, "<br>", "Number of Hispanic or Latino People: ", race_merged$estimate)
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = race_merged$estimate
)

map3<-leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = race_merged, 
              fillColor = ~pal(estimate), 
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7, 
              weight = 1, 
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal, 
            values = race_merged$estimate, 
            position = "bottomright", 
            title = "Number of Hispanic or Latino People") 
map3
```


```{r pop}
readRenviron("~/.Renviron")
options(tigris_use_cache = TRUE)

# Get land area for tracts to calculate population density
sepabg <- tracts(state = "42", county = c("101", "017", "029", "045",
                                                "091"))
st_geometry(sepabg) <- NULL
head(sepabg)
sepabg$A_KM2 <- sepabg$ALAND*1e-6

sf1_raw <- get_acs(geography = "tract", year = 2019, 
                         variables = c("B02001_001",  # Race total
                                       "B02001_002",   # Total Non-Hispanic White
                                       "B02001_003",   # Total Black population
                                       "B02001_004",   # Total Native American population 
                                       "B02001_005",   # Total Asian
                                       "B03002_001",  # Total Ethnicity
                                       "B03002_012"   # Total Hispanic or Latinx
                                       ),  
                         state = "PA", county = c("Philadelphia", "Bucks", 
                                                  "Chester", "Delaware", 
                                                  "Montgomery"))
sf1 <- sf1_raw %>%
  group_by(GEOID) %>%
  summarise(race_tot = estimate[variable == "B02001_001"],
            white_count = estimate[variable == "B02001_002"],
            black_count = estimate[variable == "B02001_003"],
            asian_count = estimate[variable == "B02001_004"],
            native_count = estimate[variable == "B02001_005"],
            hispa_tot = estimate[variable == "B03002_001"],
            hispa_count = estimate[variable == "B03002_012"]
            ) %>%
  mutate(white_per = white_count/race_tot*100,
         black_per = black_count/race_tot*100,
         asian_per = asian_count/race_tot*100,
         native_per = native_count/race_tot*100,
         hispa_per = hispa_count/hispa_tot*100,
         pop_count = race_tot) %>%
  dplyr::select(GEOID, pop_count, white_per, black_per, asian_per, native_per, hispa_per)
knitr::kable(head(sf1), "simple")
```

```{r geoid code}

append_geoid <- function(address, geoid_type = 'tract') {

  if ("lat" %in% colnames(address) && "lon" %in% colnames(address)) {
    # Call for each row of the data
    geoids <- vector(mode="character", length = nrow(address))
    for (i in 1:nrow(address)) {
      geoids[i] <- call_geolocator_latlon(address$lat[i], address$lon[i])
    }
  } else {
    # If street, city, or state columns are factors, convert them
    # Call for each row of the data
    geoids <- vector(mode="character", length = nrow(address))
    for (i in 1:nrow(address)) {
      geoids[i] <- call_geolocator(
        as.character(address$street[i]),
        as.character(address$city[i]),
        as.character(address$state[i])
      )
    }
  }
  geoids
  # Append onto database
  address <- dplyr::mutate(address, geoid = geoids)

  # AABBBCCCCCCDEEE
  if (geoid_type == 'county') {
    end <- 5
  } else if (geoid_type == 'tract') {
    end <- 11
  } else if (geoid_type == 'block group') {
    end <- 12
  } else {
    end <- 15
  }
  address <- dplyr::mutate(address,
                           geoid = ifelse(is.na(.data$geoid), NA_character_, substr(.data$geoid, 1, end)))

  return(address)
}
```

```{r correlation plot}
final <- merge(sf1, census_final, by = "GEOID")
final[] <- lapply(final, function(x) {
    if(is.factor(x)) as.numeric(as.character(x)) else x
})
sapply(final, class)
final <- final %>% dplyr::select(-GEOID, -pm10)
x <- cor(final, use = "complete.obs")
col <- colorRampPalette(c("darkorange", "white", "steelblue"))(20)
corrplot(x, type = "upper", order = "hclust", col = col)
final
```

```{r sig value plot, error=TRUE}
cor_5 <- rcorr(as.matrix(final))
M <- cor_5$r
p_mat <- cor_5$P
corrplot(M, type = "upper", order = "hclust", 
         p.mat = p_mat, sig.level = 0.01)
```
```{r coefficient plot}
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(M, method = "color", col = col(200),  
         type = "upper", insig = "blank", order = "hclust", # Add coefficient of correlation
         tl.col = "darkblue", tl.srt = 45, #Text label color and rotation
         # Combine with significance level
         p.mat = p_mat, sig.level = 0.01,  
         # hide correlation coefficient on the principal diagonal
         diag = FALSE 
         )

```
```{r violin plot}
viol <- final %>%
 pivot_longer(
   cols = ends_with("per"),
   names_to = "ethnicity",
   values_to = "percentage",
   values_drop_na = TRUE
 )

viol <- subset(viol, select = -pop_count)
viol$ethnicity <- as.factor(viol$ethnicity)
library(ggplot2)
# Basic violin plot
p <- ggplot(viol, aes(x=ethnicity, y=percentage)) + 
  geom_violin(trim = TRUE)
p

```


```{r corrplot 2}
corrplot(M, p.mat = p_mat, method = 'circle', type = 'lower', insig='blank',
         addCoef.col ='black', number.cex = 0.6, order = 'AOE', diag=FALSE)


```

```{r income}
# Get land area for tracts to calculate population density
sepabg <- tracts(state = "42", county = c("101", "017", "029", "045",
                                                "091"))
st_geometry(sepabg) <- NULL
head(sepabg)
sepabg$A_KM2 <- sepabg$ALAND*1e-6

sf2_raw <- get_acs(geography = "tract", year = 2019, 
                         variables = c("B19001_001",  # Total households
                                       "B19001_002",   # Total <$10,000
                                       "B19001_006",   # Total 25000-29999
                                       "B19001_010",   # Total 45000-49999
                                       "B19001_013",   # Total 75000-99999
                                       "B19001_015",  # Total 125000-149999
                                       "B19001_017"   # Total >200000
                                       ),  
                         state = "PA", county = c("Philadelphia", "Bucks", 
                                                  "Chester", "Delaware", 
                                                  "Montgomery"))
sf2 <- sf2_raw %>%
  group_by(GEOID) %>%
  summarise(house_tot = estimate[variable == "B19001_001"],
            pov_count = estimate[variable == "B19001_002"],
            low_count = estimate[variable == "B19001_006"],
            lower_med_count = estimate[variable == "B19001_010"],
            upper_med_count = estimate[variable == "B19001_013"],
            upper_count = estimate[variable == "B19001_015"],
            v_upper_count = estimate[variable == "B19001_017"]
            ) %>%
  mutate(pov_per = pov_count/house_tot*100,
         low_per = low_count/house_tot*100,
         lower_med_per = lower_med_count/house_tot*100,
         upper_med_per = upper_med_count/house_tot*100,
         upper_per = upper_count/house_tot*100,
         v_upper_per = v_upper_count/house_tot*100) %>%
  dplyr::select(GEOID, pov_per, low_per, lower_med_per, upper_med_per, upper_per, v_upper_per)


```

```{r correlation plot 2}
final <- merge(sf2, census_final, by = "GEOID")
final[] <- lapply(final, function(x) {
    if(is.factor(x)) as.numeric(as.character(x)) else x
})
sapply(final, class)
final <- final %>% dplyr::select(-GEOID, -pm10)
x <- cor(final, use = "complete.obs")
col <- colorRampPalette(c("darkorange", "white", "steelblue"))(20)
corrplot(x, type = "upper", order = "hclust", col = col)
```
